MINOS Handover Plots

### 2024-05-29 minos/validation/handovers.Rmd
Additions to handovers and validation

library(tidyverse)
library(ggplot2)
library(ggridges) # for densities over time plots. 
library(knitr)
library(here)
library(dplyr)
library(tidyr)
library(purrr)
library(gghighlight)
library(spatstat)

@@

```{r}
# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))
raw.dat <- raw.dat %>%
  filter(weight > 0)

out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE, 
                                    drop.zero.weight = TRUE)
#base.dat <- read_singular_local_out(out.path, 'goodHeatingDummy', drop.dead = TRUE)
```

@@

combined <- rbind(raw.household_count_by_year, base.household_count_by_year)

@@

```{r}
# Step 1: Filter for households with children
households_with_kids <- raw.dat %>%
  filter(nkids > 0) %>%
  select(hidp, time, hh_income) %>%
  distinct() %>%
  select(-hidp)

# Step 2: Calculate income quintiles
households_with_kids <- households_with_kids %>%
  group_by(time) %>%
  mutate(income_quintile = ntile(hh_income, 5)) %>%
  ungroup()

# Step 3: Count households within each income quintile per year
household_count_by_income_quintile <- households_with_kids %>%
  group_by(time, income_quintile) %>%
  summarise(n = n())

household_count_by_income_quintile
```

@@

### Poverty Thresholds

#### Relative
```{r}
rel.threshold.raw <- raw.dat %>%
  select(time, weight, hh_income) %>%
  group_by(time) %>%
  summarise(poverty_threshold = (weighted.mean(hh_income, w = weight) * 0.6))

rel.threshold.base <- base.dat %>%
  select(time, weight, hh_income) %>%
  group_by(time) %>%
  summarise(poverty_threshold = (weighted.mean(hh_income, w = weight) * 0.6)) %>%
  filter(time != 2021)

rel.threshold <- rbind(rel.threshold.raw, rel.threshold.base)

ggplot(data = rel.threshold, aes(x = time, y = poverty_threshold)) +
  geom_col() +
  geom_hline(yintercept = rel.threshold$poverty_threshold[rel.threshold$time == 2020], linetype='dotted') +
  geom_vline(xintercept = start.year, linetype = 'dashed') +
  labs(title = 'Relative poverty threshold over time') +
  xlab('Year') +
  ylab('Threshold')
#rm(rel.threshold)
```

#### Absolute
```{r}
raw.2010 <- raw.dat %>%
  filter(time == 2010)

abs.threshold <- weighted.median(x = raw.2010$hh_income, 
                                 w = raw.2010$weight, 
                                 na.rm = TRUE)
abs.threshold <- abs.threshold * 0.6

print(sprintf("The absolute poverty threshold is £%.02f", abs.threshold))
```

@@

## S7 Labour State

```{r}
prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'S7_labour_state')

lab.state <- prepped.list$var
lab.state.norm <- prepped.list$norm

# Now plot
ggplot(data = lab.state, mapping = aes(x = time, y = n, group = S7_labour_state, colour = S7_labour_state)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'S7 Labour State', subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = lab.state.norm, mapping = aes(x = time, y = prct, fill=S7_labour_state)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'S7 Labour State') +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'S7_labour_state.png',
         plot = last_plot(),
         path = save.path)
}

rm(prepped.list, lab.state, lab.state.norm)
```

```{r}
pivoted <- combine_and_pivot_long(df1 = base.dat, 
                                       df1.name = 'simulated', 
                                       df2 = raw.dat, 
                                       df2.name = 'raw', 
                                       var = 'S7_labour_state')

cv_ordinal_plots(pivoted.df = pivoted, 
                 var = 'S7_labour_state',
                 save = shall.we.save,
                 save.path = save.path)
```

# ADDITIONAL VARIABLES

@@

## Behind On Bills

```{r}
raw.behind_on_bills <- raw.dat %>%
  dplyr::select(pidp, time, behind_on_bills) %>%
  group_by(time, behind_on_bills) %>%
  count() %>%
  mutate(source = 'final_US')

base.behind_on_bills <- base.dat %>%
  dplyr::select(pidp, time, behind_on_bills) %>%
  group_by(time, behind_on_bills) %>%
  count() %>%
  mutate(source = 'baseline_output')

# merge before plot
behind_on_bills <- rbind(raw.behind_on_bills, base.behind_on_bills)
behind_on_bills$behind_on_bills <- as.factor(behind_on_bills$behind_on_bills)

# Now plot
ggplot(data = behind_on_bills, mapping = aes(x = time, y = n, group = behind_on_bills, colour = behind_on_bills)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Behind On Bills', subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

behind_on_bills.norm <- behind_on_bills %>%
  group_by(time) %>%
  mutate(total = sum(n)) %>%
  mutate(prct = (n / total))

ggplot(data = behind_on_bills.norm, mapping = aes(x = time, y = prct, fill=behind_on_bills)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Behind On Bills') +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'behind_on_bills.png',
       plot = last_plot(),
       path = save.path)
}

rm(raw.behind_on_bills, base.behind_on_bills, behind_on_bills, behind_on_bills.norm)
```

## nkids Household

```{r}
raw.hh_unique <- raw.dat %>%
  select(hidp, time, nkids) %>%
  distinct()

base.hh_unique <- base.dat %>%
  select(hidp, time, nkids) %>%
  distinct()

#handover_boxplots(raw.hh_unique, base.hh_unique, 'nkids')
handover_lineplots(raw.hh_unique, base.hh_unique, 'nkids')
```

```{r}
raw.s <- raw.dat %>%
  select(pidp, age, time, nkids)
base.s <- base.dat %>%
  select(pidp, age, time, nkids)

spaghetti_plot(spag, 'nkids',
               save = shall.we.save,
               save.path = save.path)

## nkids Max

```{r}
# first figure out how to plot final_US
# start with hh_income
raw.nkids_max <- raw.dat %>%
  dplyr::select(pidp, time, nkids) %>%
  group_by(time) %>%
  summarise(max_nkids = max(nkids)) %>%
  mutate(source = 'final_US')

base.nkids_max <- base.dat %>%
  dplyr::select(pidp, time, nkids) %>%
  group_by(time) %>%
  summarise(max_nkids = max(nkids)) %>%
  mutate(source = 'baseline_output')

# merge before plot
max_nkids <- rbind(raw.nkids_max, base.nkids_max)

# Now plot
ggplot(data = max_nkids, mapping = aes(x = time, y = max_nkids, group = source, colour = source)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Max nkids', subtitle = 'Full Sample') +
  xlab('Year') +
  ylab('Average')
```

@@

## Hourly Wage

```{r}
handover_boxplots(raw.dat, base.dat, 'hourly_wage')
handover_lineplots(raw.dat, base.dat, 'hourly_wage')
```

```{r}
raw.s <- raw.dat %>%
  select(pidp, age, time, hourly_wage)
base.s <- base.dat %>%
  select(pidp, age, time, hourly_wage)

spaghetti_plot(spag, 'hourly_wage',
               save = shall.we.save,
               save.path = save.path)

# now a version limited to 0-100
spag <- spag %>%
  filter(hourly_wage <= 300)
spaghetti_plot(spag, 'hourly_wage',
               save = shall.we.save,
               save.path = save.path)

@@

```{r}
raw.child.ages <- raw.dat %>%
  dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>%
  filter(time > 2013 & child_ages != 'childless') %>%
  group_by(time, hidp) %>%
  summarise(child_ages = find_mode(child_ages))

r.s<- strsplit(raw.child.ages$child_ages, "_")

raw.child.ages.expanded <- data.frame(time = rep(raw.child.ages$time, sapply(r.s, length)), expanded_child_ages = unlist(r.s))
raw.child.ages.expanded <- raw.child.ages.expanded %>%
  group_by(time, expanded_child_ages) %>%
  summarise(age_counts = n()) %>%
  mutate(source = 'final_US') %>%
  filter(expanded_child_ages <= 16)

base.child.ages <- base.dat %>%
  dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>%
  filter(time > 2009) %>%
  group_by(time, hidp) %>%
  summarise(child_ages = find_mode(child_ages))

b.s<- strsplit(base.child.ages$child_ages, "_")

base.child.ages.expanded <- data.frame(time = rep(base.child.ages$time, sapply(b.s, length)), expanded_child_ages = unlist(b.s))
base.child.ages.expanded <- base.child.ages.expanded %>%
  group_by(time, expanded_child_ages) %>%
  summarise(age_counts = n()) %>%
  mutate(source = 'baseline')

combined <- rbind(raw.child.ages.expanded, base.child.ages.expanded)

combined$expanded_child_ages <- factor(combined$expanded_child_ages,
                                       levels = c('0', '1', '2', '3', '4', '5', 
                                                  '6', '7', '8', '9', '10', 
                                                  '11', '12', '13', '14', '15', 
                                                  '16', 'childless'))

ggplot(data = combined, mapping = aes(x = time, y = age_counts, group = expanded_child_ages, colour =  factor(as.numeric(expanded_child_ages)))) +
  geom_line() +
  geom_vline(xintercept = start.year) +
  labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') +
  scale_colour_viridis_d() +
  xlab('Year') +
  ylab('Count')

ggplot(data = combined, mapping = aes(x = expanded_child_ages, y = age_counts, group = factor(time), colour = factor(time))) +
  geom_line() +
  geom_vline(xintercept = start.year) +
  labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') +
  scale_colour_viridis_d() +
  xlab('Age') +
  ylab('Count')

base.child.ages.nkids <- base.dat %>%
  group_by(time, hidp) %>%
  summarise(nkids_counts = mean(nkids)) %>%
  group_by(time) %>%
  summarise(nkids_by_year = sum(nkids_counts)) %>%
  mutate(source = 'baseline')

raw.child.ages.nkids <- raw.dat %>%
  group_by(time, hidp) %>%
  summarise(nkids_counts = mean(nkids)) %>%
  group_by(time) %>%
  summarise(nkids_by_year = sum(nkids_counts)) %>%
  mutate(source = 'final_US')

combined.nkids <- rbind(base.child.ages.nkids, raw.child.ages.nkids)

ggplot(data = combined.nkids, mapping = aes(x = time, y = nkids_by_year, group = source, colour = source)) +
  geom_line() +
  labs(title = "nkids Count Over Time", subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

@@

```{r}

@@

>>>>

require(tidyverse)
require(ggplot2)
require(ggridges) # for densities over time plots. 
require(knitr)
require(here)
require(dplyr)
require(tidyr)
require(purrr)
require(gghighlight)

@@

```{r}
# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)

#out.path <- here::here('output', 'scotland_scaled/')
#base.dat <- read_first_singular_local_out(out.path, '25UniversalCredit', drop.dead = TRUE)

# HR 283-285 older version (285)
# out.path <- here::here('output', 'child_poverty_config/')
# base.dat <- read_singular_local_out(out.path, 'ChildPovertyIntervention', drop.dead = TRUE)
```

@@

combined <- rbind(raw.households_with_kids, base.households_with_kids)

@@

@@

@@

@@

@@

@@

```{r}
base.child.ages <- base.dat %>%
  dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>%
  filter(time > 2009) %>%
  group_by(time, hidp) %>%
  summarise(child_ages = find_mode(child_ages))

s<- strsplit(base.child.ages$child_ages, "_")

base.child.ages.expanded <- data.frame(time = rep(base.child.ages$time, sapply(s, length)), expanded_child_ages = unlist(s))
base.child.ages.expanded <- base.child.ages.expanded %>%
  group_by(time, expanded_child_ages) %>%
  summarise(age_counts = n())

ggplot(data = base.child.ages.expanded, mapping = aes(x = time, y = age_counts, group = expanded_child_ages, colour =  factor(as.numeric(expanded_child_ages)))) +
  geom_line() +
  labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') +
  scale_colour_viridis_d() +
  xlab('Year') +
  ylab('Count')

ggplot(data = base.child.ages.expanded, mapping = aes(x = factor(as.numeric(expanded_child_ages)), y = age_counts, group = factor(time), colour = factor(time))) +
  geom_line() +
  labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') +
  scale_colour_viridis_d() +
  xlab('Year') +
  ylab('Count')

base.child.ages.nkids <- base.dat %>%
  group_by(time, hidp) %>%
  summarise(nkids_counts = mean(nkids)) %>%
  group_by(time) %>%
  summarise(nkids_by_year = sum(nkids_counts))

ggplot(data = base.child.ages.nkids, mapping = aes(x = time, y = nkids_by_year)) +
  geom_line() +
  labs(title = "nkids Count Over Time", subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

@@

## CHILD POVERTY VARS

### Relative poverty

```{r}
prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'relative_poverty')

tit <- 'Relative Poverty'

# Now plot
ggplot(data = var, mapping = aes(x = time, y = n, group = relative_poverty, colour = relative_poverty)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit, subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = norm, mapping = aes(x = time, y = prct, fill=relative_poverty)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit) +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'relative_poverty.png',
         plot = last_plot(),
         path = save.path)
}

### Absolute Poverty

```{r}
prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'absolute_poverty')

tit <- 'Absolute Poverty'

# Now plot
ggplot(data = var, mapping = aes(x = time, y = n, group = absolute_poverty, colour = absolute_poverty)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit, subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = norm, mapping = aes(x = time, y = prct, fill=absolute_poverty)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit) +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'absolute_poverty.png',
         plot = last_plot(),
         path = save.path)
}

### Persistent Poverty

```{r}
prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'persistent_poverty')

tit <- 'Persistent Poverty'

# Now plot
ggplot(data = var, mapping = aes(x = time, y = n, group = persistent_poverty, colour = persistent_poverty)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit, subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = norm, mapping = aes(x = time, y = prct, fill=persistent_poverty)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit) +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'persistent_poverty.png',
         plot = last_plot(),
         path = save.path)
}

rm(prepped.list, var, norm, tit)

@@

### Low Income and Material Deprivation

```{r}
prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'low_income_matdep_child')

tit <- 'Material Deprivation Child'

# Now plot
ggplot(data = var, mapping = aes(x = time, y = n, group = low_income_matdep_child, colour = low_income_matdep_child)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit, subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = norm, mapping = aes(x = time, y = prct, fill=low_income_matdep_child)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = tit) +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'low_income_matdep_child.png',
         plot = last_plot(),
         path = save.path)
}

rm(prepped.list, var, norm, tit)
